Classless Java: Tuning Java Interfaces

### 2015-11-30 paper/sections/CaseStudy.tex
compressing text

The \emph{Expression Problem} (EP)~\cite{wadler98expression}
is a well-known problem about modular extensibility issues in software
evolution. Recently, a new solution~\cite{eptrivially}, \marco{repharesed} using only
covariant type refinement, has been shown.
When such solution is modelled with interfaces and default methods, it can
even provide independent extensibility: the ability to
assemble a system from multiple, independently developed, extensions.
Unfortunatelly, the required instantiation code makes a plain Java solution verbose and cumbersome to use. The \mixin annotation is enough to remove the boilerplate code, making the presented approach very appealing.
% \bruno{We should try to do some space saving in here. You could, for example 
% combine the code for the initial system with the code for a new operation in a figure/table 
% with 2 colums. Then just refer to the Table/Figure in the text.}

\paragraph{Initial System}
\yanlin{code figure updated.}
In the formulation of the EP there is an initial system that models
arithmetic expressions with only literals and addition, and an initial 
operation \texttt{eval} for expression evaluation.

@@

\noindent We use embedded DSLs of two simple SQL query languages to illustrate.
The first query language \texttt{Database}  models
\texttt{select}, \texttt{from} and \texttt{where} clauses:
\lstinputlisting[linerange=6-11]{../UseMixinLombok/src/jooq/DatabaseTest.java}% APPLY:linerange=FLUENT_DATABASE
% \bruno{Wouldn't it be better to have a static method create() that calls of("","","")) and initializes all the fields? It seems awkward 
% to have to use the of explicitly to initialized all fields.}

@@

\lstinputlisting[linerange=33-33]{../UseMixinLombok/src/jooq/DatabaseTest.java}% APPLY:linerange=FLUENT_QUERY1

@@

\paragraph{Extending the Query Language} The previous query language can be extended with a new feature
\texttt{orderBy} which orders the result records by a field that users
specify. With \mixin programmers just need to extend the interface \texttt{Database} with new
feature, and the return type of fluent setters in
\texttt{Database} is automatically refined to \texttt{ExtendedDatabase}:
\lstinputlisting[linerange=15-18]{../UseMixinLombok/src/jooq/DatabaseTest.java}% APPLY:linerange=FLUENT_DATABASE_EXT
This way, when a \texttt{ExtendedDatabase} query created, 
all the fluent setters return the correct type, and not the old \texttt{Database} type, which would prevent calling 
\texttt{orderBy}.

\lstinputlisting[linerange=39-39]{../UseMixinLombok/src/jooq/DatabaseTest.java}% APPLY:linerange=FLUENT_QUERY2

>>>>

The \emph{Expression Problem} (EP)~\cite{wadler98expression} is a well-known
problem about modular extensibility issues in software evolution. Recently, a
new solution~\cite{eptrivially}, using only covariant type refinement, has been
shown.  When such solution is modelled with interfaces and default methods, it
can even provide independent extensibility: the ability to assemble a system
from multiple, independently developed, extensions.  Unfortunatelly, the
required instantiation code makes a plain Java solution verbose and cumbersome
to use. The \mixin annotation is enough to remove the boilerplate code, making
the presented approach very appealing.

\paragraph{Initial System}
In the formulation of the EP there is an initial system that models
arithmetic expressions with only literals and addition, and an initial 
operation \texttt{eval} for expression evaluation.

@@

\noindent We use embedded DSLs of two simple SQL query languages to illustrate.
The first query language \texttt{Database}  models
\texttt{select}, \texttt{from} and \texttt{where} clauses:
\lstinputlisting[linerange=6-10]{../UseMixinLombok/src/jooq/DatabaseTest.java}% APPLY:linerange=FLUENT_DATABASE
% \bruno{Wouldn't it be better to have a static method create() that calls of("","","")) and initializes all the fields? It seems awkward 
% to have to use the of explicitly to initialized all fields.}

@@

\lstinputlisting[linerange=31-31]{../UseMixinLombok/src/jooq/DatabaseTest.java}% APPLY:linerange=FLUENT_QUERY1

@@

\paragraph{Extending the Query Language} The previous query language can be extended with a new feature
\texttt{orderBy} which orders the result records by a field that users
specify. With \mixin programmers just need to extend the interface \texttt{Database} with new
feature, and the return type of fluent setters in
\texttt{Database} is automatically refined to \texttt{ExtendedDatabase}:
\lstinputlisting[linerange=14-16]{../UseMixinLombok/src/jooq/DatabaseTest.java}% APPLY:linerange=FLUENT_DATABASE_EXT
This way, when a \texttt{ExtendedDatabase} query created, 
all the fluent setters return the correct type, and not the old \texttt{Database} type, which would prevent calling 
\texttt{orderBy}.

\lstinputlisting[linerange=37-37]{../UseMixinLombok/src/jooq/DatabaseTest.java}% APPLY:linerange=FLUENT_QUERY2

### 2015-11-30 paper/sections/Introduction.tex
compressing text

To evaluate the usefulness of object interfaces, we illustrate 3
applications and case studies. The first application is a simple 
solution to the Expression Problem~\cite{wadler98expression}, supporting independent 
extensibility~\cite{zenger05independentlyextensible}, and without boilerplate code. The second
application is to show how embedded DSLs using fluent interfaces~\cite{fowler2005fluentinterface} 
can be easily defined using object interfaces. Finally, the last
application is a larger case study for a simple Maze game implemented with 
multiple inheritance. For the last application we show that there is a
significant reduction in the numbers of lines of code when compared 
to an existing implementation~\cite{bono14} using plain Java 8. 
Noteworthy, is the fact that all applications are implemented 
without defining a single class!

In summary, the contributions of this paper are:

>>>>

To evaluate the usefulness of object interfaces, we illustrate 3
applications (case studies). The first application is a simple 
solution to the Expression Problem~\cite{wadler98expression}, supporting independent 
extensibility~\cite{zenger05independentlyextensible}, and without boilerplate code. The second
application is shows how embedded DSLs using fluent interfaces~\cite{fowler2005fluentinterface} 
can be easily defined using object interfaces. The last
application is a larger case study for a simple Maze game implemented with 
multiple inheritance. For the last application we show that there is a
significant reduction in the numbers of lines of code when compared 
to an existing implementation~\cite{bono14} using plain Java 8. 
Noteworthy, all applications are implemented 
without defining a single class!

In summary, the contributions of this paper are:
\begin{itemize}

### 2015-11-30 paper/sections/Overview.tex
code for expression problem updated.
compressing text

\paragraph{Instantiation}
To use \texttt{Horse}, \texttt{Bird} and \texttt{Pegasus}, some
objects must be instantiated first. A first problem with using
interfaces to model the animal system is simply that interfaces 
cannot directly be instantiated: a class is needed. Therefore, we have
to first create some classes, such as:

@@

\noindent in order to create animal objects. With those classes, 
a Pegasus animal, can be created using the class constructor:

@@

\noindent There are a couple of annoyances here. Firstly, the sole
purpose of the classes is to provide a way to instantiate the
objects. Although (in this case) it takes only one line code to
provide each of those classes, this code is essentially boilerplate
code, which does not add behaviour to the system. Secondly, 
the namespace gets filled with three additional types. For example, 
both \texttt{Horse} and \texttt{HorseImpl} are needed: \texttt{Horse}
is needed because it needs to be an inferface so that \texttt{Pegasus}
can use multiple inheritance; and \texttt{HorseImpl} is needed to
provide object instantiation. 
Note that, for this very simple animal system, plain Java 8 anonymous
classes can be used to avoid these problems.  We could have simply
instantiated Pegasus using:

@@

\begin{lstlisting}
interface Pegasus extends Horse, Bird { // generated code not visible to users
	static Pegasus of() {return new Pegasus() {};}
}
\end{lstlisting}

@@

\paragraph{\Q@Point2D@: simple immutable data with fields}
Two dimensional points should keep track of their coordinates. 
The usual approach to model points in Java would be to use a class 
with fields for the coordinates. However here we will illustrate how 
points are modelled with interfaces, providing a simple 
example of how to model immutable state with interfaces. Since Java disallows fields 
inside interfaces, the state is modelled using abstract (getter) methods:

@@

\noindent Unfortunatelly, creating a new point object is quite cumbersome, even 
when using anonymous classes:

@@

\noindent However programmers are not required to use this cumbersome syntax for
every object allocation. As programmers do, for ease or reuse, the
boring long repetitive code can be encapsulate in a method. A
generalization of the \texttt{of} static factory
method is appropriate in this case:
\begin{lstlisting}
interface Point2D{ int x(); int y();
  static Point2D of(int x, int y){return new Point2D(){
    public int x(){return x;} public int y(){return y;}};}
}
\end{lstlisting}

\vspace{-5pt}
\paragraph{\Q@Point2D@ with Object Interfaces}
This obvious constructor code can be automatically generated by our
\mixin annotation.  By annotating the interface \Q@Point2D@, the
annotation will generate a variation of the shown static method
\texttt{of}, mimicking the functionality of a simple minded
constructor: By looking at the methods that need implementations it
first detects what are the fields, then generates an \Q@of@ method with
one argument for each of them. That is, we can just write

@@

%Now we proceed to define \texttt{Animal} with \texttt{point} ``member
%field''.
%%Not again, before we used only getters!
% Again, we model this member field with getter and setter methods:
\lstinputlisting[linerange=59-62]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=ANIMAL

@@

\lstinputlisting[linerange=66-69]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=HORSE

@@

\subsection{Object Interfaces and Subtyping}

%\subsection{\Q@Bird@:  field type refinement}

\Q@Bird@s are \Q@Animal@s, but while \Q@Animal@s only need 2D
locations, \Q@Bird@s need 3D locations. Therefore when the \texttt{Bird}
interface extends the animal interface, the notion of points needs to
be refined. Such kind of refinement usually poses a challenge
with typical class-based approaches. Fortunatelly, with object interfaces,
we are able to provide a simple and effective solution to that problem.

@@

\lstinputlisting[linerange=73-81]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=BIRD

@@

\paragraph{Generated Boilerplate}
Just to give a feeling on how much boring code \mixin is generating,
we show the generated code for the \texttt{Point3D} interface 
Figure~\ref{fig:boilerplate}. The generated code is very repetitive,
writing such code by hand can easily induce bugs, for example a
distracted programmer may swap the arguments of one of the many calls of
\Q@Point3D.of@.
\begin{figure}
\begin{lstlisting}
interface Point3D extends Point2D{
  Point3D withX(int val);    Point3D withY(int val);    Point3D withZ(int val);
  Point3D with(Point2D val);
  public static Point3D of(int _x, int _y, int _z){
    int x=_x;    int y=_y;    int z=_z;
    return new Point3D(){
      public int x(){return x;}    public int y(){return y;}    public int z(){return z;}
      public Point3D withX(int val){return Point3D.of(val,this.y(),this.z());}
      public Point3D withY(int val){return Point3D.of(this.x(),val,this.z());}
      public Point3D withZ(int val){return Point3D.of(this.x(),this.y(),val);}
      public Point3D with(Point2D val){
        if(val instanceof Point3D){return (Point3D)val;}
        return Point3D.of(val.x(),val.y(),this.z());
        }      };    }  }
\end{lstlisting}
\caption{Generated boilerplate code.}
\label{fig:boilerplate}
\end{figure}
Note how \Q@with-@ methods are automatically refined in their return type, so that code like 
\Q@Point3D p=Point3D.of(1,2,3); p=p.withX(42);@ will be accepted and behave as expected. If the programmer wishes to suppress this behaviour and keep the signature as it was, it is sufficient to redefine the \Q@with-@ methods in the new class repeating the old signature.
Again, the philosophy is that if the programmer provides something directly, \mixin does not touch it.
\Q@with-@ methods (functionally) update a field/property at a time.
This can be inefficient, and sometime hard to maintain.
Often we want to update many fields at the same time, for example using another object as source.
Following this idea, the method \Q@with(Point2D)@ is an example of a (functional) properties updater: 
it takes a certain type and in the current object update all field in that type that match fields in the current type.
The idea is that we want as result something that is still like \Q@this@, but modified to be as much as possible similar to the parameter.
The cast in \Q@with(Point2D)@ is trivially safe since it is guarded by an \Q@instanceof@ test. The idea is that if the parameter is a subtype of the current exact type, then we can just return the parameter, as something that is just ``more'' than \Q@this@.

@@

\lstinputlisting[linerange=85-86]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=PEGASUS

@@

\lstinputlisting[linerange=51-55]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT3D

\lstinputlisting[linerange=59-62]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=ANIMAL

\lstinputlisting[linerange=66-69]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=HORSE

\lstinputlisting[linerange=73-81]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=BIRD

\lstinputlisting[linerange=85-86]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=PEGASUS

>>>>

\paragraph{Instantiation}
To use \texttt{Horse}, \texttt{Bird} and \texttt{Pegasus}, some
objects must be created first. A first problem with using
interfaces to model the animal system is simply that interfaces 
cannot be directly instantiated. Classes are needed for instantiation, such as:

@@

\noindent to create animal objects. Now a \texttt{Pegasus} animal can be created
using the class constructor:

@@

\noindent There is a lot of annoyance here. Firstly, the sole
purpose of classes is to provide a way to instantiate
objects. Although (in this case) it takes only one line code to
provide each of those classes, this code is essentially boilerplate
code, which does not add behaviour to the system. Secondly, 
the namespace gets filled with three additional types. For example, 
both \texttt{Horse} and \texttt{HorseImpl} are needed: \texttt{Horse}
is needed because it needs to be an inferface so that \texttt{Pegasus}
can use multiple inheritance; and \texttt{HorseImpl} is needed to
provide object instantiation. 
Note that, for this very simple animal system, plain Java 8 anonymous
classes can be used to avoid these problems.  We could have simply
instantiated Pegasus using:

@@

\begin{lstlisting}
interface Pegasus extends Horse, Bird { // generated code not visible to users
	static Pegasus of() {return new Pegasus() {};}}
\end{lstlisting}

@@

\paragraph{\Q@Point2D@: simple immutable data with fields}
Two dimensional points should keep track of their coordinates.  The usual
approach to model points in Java would be to use a class with fields for the
coordinates. However here we will illustrate how points are modelled with
interfaces. Since Java disallows fields inside interfaces, states are modeled
using abstract (getter) methods:

@@

\noindent Unfortunatelly, creating a new point object is quite cumbersome, even
with anonymous classes:

@@

\noindent However it is not required to use this cumbersome syntax for every
object allocation. As programmers do, for ease or reuse, the boring
repetitive code can be encapsulated in a method. A generalization of the
\texttt{of} static factory method is appropriate in this case:
\begin{lstlisting}
interface Point2D{ int x(); int y();
  static Point2D of(int x, int y){return new Point2D(){
    public int x(){return x;} public int y(){return y;}};}}
\end{lstlisting}

\vspace{-5pt}
\paragraph{\Q@Point2D@ with Object Interfaces}
This obvious constructor code can be automatically generated by our
\mixin annotation.  By annotating the interface \Q@Point2D@, a variation of the shown static method
\texttt{of} will be generated, mimicking the functionality of a simple-minded
constructor: by looking at the methods that need implementation it
first detects what are the fields, then generates an \Q@of@ method with
one parameter for each of them. That is, we can just write

@@

%Now we proceed to define \texttt{Animal} with \texttt{point} ``member
%field''.
%%Not again, before we used only getters!
% Again, we model this member field with getter and setter methods:
\lstinputlisting[linerange=59-59]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=ANIMAL

@@

\lstinputlisting[linerange=63-64]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=HORSE

@@

\subsection{Object Interfaces and Subtyping}
\Q@Bird@s are \Q@Animal@s, but while \Q@Animal@s only need 2D
locations, \Q@Bird@s need 3D locations. Therefore when the \texttt{Bird}
interface extends the animal interface, the notion of points needs to
be refined. Such kind of refinement usually poses a challenge
in typical class-based approaches. Fortunatelly, with object interfaces,
we are able to provide a simple and effective solution to that problem.

@@

\lstinputlisting[linerange=68-72]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=BIRD

@@

\paragraph{Generated boilerplate}
Just to give a feeling on how much boring code \mixin is generating,
we show the generated code for the \texttt{Point3D} interface 
Figure~\ref{fig:boilerplate}. The generated code is very repetitive,
writing such code by hand can easily induce bugs, for example a
distracted programmer may swap the arguments of one of the many calls of
\Q@Point3D.of@.
\begin{figure}
\begin{lstlisting}
interface Point3D extends Point2D{
    Point3D withX(int val); Point3D withY(int val); Point3D withZ(int val);
    Point3D with(Point2D val);
    public static Point3D of(int _x, int _y, int _z){
        int x=_x;    int y=_y;    int z=_z;
        return new Point3D(){
            public int x(){return x;}    public int y(){return y;}    public int z(){return z;}
            public Point3D withX(int val){return Point3D.of(val,this.y(),this.z());}
            public Point3D withY(int val){return Point3D.of(this.x(),val,this.z());}
            public Point3D withZ(int val){return Point3D.of(this.x(),this.y(),val);}
            public Point3D with(Point2D val){
                if(val instanceof Point3D){return (Point3D)val;}
                return Point3D.of(val.x(),val.y(),this.z()); }};}}
\end{lstlisting}
\caption{Generated boilerplate code.}
\label{fig:boilerplate}
\end{figure}
Note how \Q@with-@ methods are automatically refined in their return type, so that code like 
\Q@Point3D p=Point3D.of(1,2,3); p=p.withX(42);@ will be accepted and behave as expected. If the programmer wishes to suppress this behaviour and keep the signature as it was, it is sufficient to redefine the \Q@with-@ methods in the new class repeating the old signature.
Again, the philosophy is that if the programmer provides something directly, \mixin does not touch it.
\Q@with-@ methods (functionally) update a field/property at a time.
This can be inefficient, and sometime hard to maintain.
Often we want to update many fields at the same time, for example using another object as source.
Following this idea, the method \Q@with(Point2D)@ is an example of a (functional) properties updater: 
it takes a certain type and in the current object update all field in that type that match fields in the current type.
The idea is that we want as result something that is still like \Q@this@, but modified to be as much as possible similar to the parameter.
The cast in \Q@with(Point2D)@ is trivially safe since it is guarded by an \Q@instanceof@ test. The idea is that if the parameter is a subtype of the current exact type, then we can just return the parameter, as something that is just ``more'' than \Q@this@.

@@

\lstinputlisting[linerange=76-76]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=PEGASUS

@@

\lstinputlisting[linerange=51-56]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT3D

\lstinputlisting[linerange=59-59]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=ANIMAL

\lstinputlisting[linerange=63-64]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=HORSE

\lstinputlisting[linerange=68-72]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=BIRD

\lstinputlisting[linerange=76-76]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=PEGASUS

### 2015-11-30 paper/sections/RelatedWork.tex
compressing text

\subsection{Multiple Inheritance in Object Oriented Languages}
Many authors have argued in favour or against multiple inheritance.  Multiple inheritance provides
expressive power, but it is difficult to model and implement, and can create
programs that are hard to reason about. These difficulties include the famous
diamond (fork-join) problem~\cite{bracha90mixin,Sak89dis}, conflicting methods, etc.
%and the yo-yo problem~\cite{taenzer1989problems}.  
To conciliate the need for expressive power and the need for simplicity, many
models have been proposed in over the years, including C++ virtual
inheritance, mixins~\cite{bracha90mixin}, traits~\cite{scharli03traits}, and
hybrid model such as CZ~\cite{malayeri2009cz}.
They provide novel programming architecture models in the OO paradigm. In terms
of restrictions set on these models, C++ virtual inheritance aims at a more
general model, mixins added some restrictions on the model, and trait model is
the most restricted one (excluding states, instantiation, etc).
\bruno{How about Malayeri and Aldrich's paper? shouldn't that be discussed?}
\yanlin{Newly added discussion on CZ. please double check.}

\paragraph{C++ Approach.}
C++ tries to provide a general solution to multiple inheritance. Virtual
inheritance in C++ provides another solution to multiple inheritance (especially
the diamond problem by keeping only one copy of the base
class)~\cite{ellis1990annotated}, however suffers from object initialization
problem as pointed out by Malayeri et al.~\cite{malayeri2009cz}. It bypasses all
constructor calls to virtual superclasses, which would potentially cause serious
semantic errors. \bruno{What happens in our approach for the same
  case?}\yanlin{explained below please double-check.} In our
approach, \mixin annotation processor has the full control of object
initialization mechanism, and this mechanism is clean and transparent to
users. Moreover, if not satisfied with the default generated \texttt{of} method,
customized factory methods are also allowed to be defined.

@@

\bruno{rrevise latter?}
However, in pure Java, there is no mechanism for creating objects in
interfaces. Also, our approach supports proper constructor mechanism.

@@

\paragraph{Traits vs Object Interfaces}
We consider object interfaces to be an alternative to traits or mixins.
In the trait model two concepts (traits and classes) coexist and
cooperate. Some authors~\cite{bettini2013pure}\yanlin{There are many trait related papers of ferruccio
  damiani, please check whether citing this one is ok.} 
see this as good language design
fostering good software development by helping programmers to think
about the structure of their programs.  However, other authors see the
need of two concepts and the absence of state as drawbacks of this
model~\cite{malayeri2009cz}.  Object interfaces are units of reuse, and at
the same time they provide factory methods for instantiation and
support state.  Our approach promotes the use of interfaces instead of
classes, in order to rely on the modular composition offered
by interfaces. Since Java was designed for classes, a direct class-less
programming style is verbose and feels unnatural. However, annotation
driven code generation is enough to overcome this difficulty, and the
resulting programming style encourages modularity, composability and
reusability, by keeping a strong object oriented feel. In that sense,
we promote object interfaces as being both units of reusable code and
object factories. Our practical experience is that, in Java,
separating the two notions leads to alot of boilerplate code, and is
quite limiting when multiple inheritance with state is required. The
use of abstract state operations avoids the key difficulties
associated with multiple inheritance and state, while still being quite expressive. 
Moreover the ability to support constructors 
adds additional expressivity, which is not available in approaches such 
as Scala's traits/mixins.

>>>>

\subsection{Multiple Inheritance in Object Oriented Languages}
Many authors have argued in favour or against multiple inheritance.  Multiple inheritance provides
expressive power, but it is difficult to model and implement, and can create
programs that are hard to reason about. These difficulties include the famous
diamond (fork-join) problem~\cite{bracha90mixin,Sak89dis}, conflicting methods, etc.
%and the yo-yo problem~\cite{taenzer1989problems}.  
To conciliate the need for expressive power and the need for simplicity, many
models have been proposed in over the years, including C++ virtual
inheritance, mixins~\cite{bracha90mixin}, traits~\cite{scharli03traits}, and
hybrid model such as CZ~\cite{malayeri2009cz}.
They provide novel programming architecture models in the OO paradigm. In terms
of restrictions set on these models, C++ virtual inheritance aims at a more
general model, mixins added some restrictions on the model, and trait model is
the most restricted one (excluding states, instantiation, etc).

\paragraph{C++ Approach.}
C++ tries to provide a general solution to multiple inheritance. Virtual
inheritance in C++ provides another solution to multiple inheritance (especially
the diamond problem by keeping only one copy of the base
class)~\cite{ellis1990annotated}, however suffers from object initialization
problem as pointed out by Malayeri et al.~\cite{malayeri2009cz}. It bypasses all
constructor calls to virtual superclasses, which would potentially cause serious
semantic errors. In our approach, \mixin annotation processor has the full
control of object initialization mechanism, and this mechanism is clean and
transparent to users. Moreover, if not satisfied with the default generated
\texttt{of} method, customized factory methods are also allowed to be defined.

@@

% However, in pure Java, there is no mechanism for creating objects in
% interfaces. Also, our approach supports proper constructor mechanism.

@@

\paragraph{Traits vs Object Interfaces}
We consider object interfaces to be an alternative to traits or mixins.  In the
trait model two concepts (traits and classes) coexist and cooperate. Some
authors~\cite{bettini2013pure} see this as good language design fostering good
software development by helping programmers to think about the structure of
their programs.  However, other authors see the need of two concepts and the
absence of state as drawbacks of this model~\cite{malayeri2009cz}.  Object
interfaces are units of reuse, and at the same time they provide factory methods
for instantiation and support state.  Our approach promotes the use of
interfaces instead of classes, in order to rely on the modular composition
offered by interfaces. Since Java was designed for classes, a direct class-less
programming style is verbose and feels unnatural. However, annotation driven
code generation is enough to overcome this difficulty, and the resulting
programming style encourages modularity, composability and reusability, by
keeping a strong object oriented feel. In that sense, we promote object
interfaces as being both units of reusable code and object factories. Our
practical experience is that, in Java, separating the two notions leads to alot
of boilerplate code, and is quite limiting when multiple inheritance with state
is required. The use of abstract state operations avoids the key difficulties
associated with multiple inheritance and state, while still being quite
expressive.  Moreover the ability to support constructors adds additional
expressivity, which is not available in approaches such as Scala's
traits/mixins.
